---
title: "Homework_10"
author: "Kyle Grasso"
date: "2024-04-03"
output: html_document
---
```{r, message=FALSE, warning=FALSE,echo = FALSE}
library(tidyverse)
library(dplyr)
library(datasets)
```

### Question 1:
```{r, message=FALSE, warning=FALSE}
# create function
count_zeros <- function(x){ 
  counter <- 0
  for (i in x) {
    if (i == 0) {
      counter <- counter + 1
    }
  }
  return(counter) 
} 

# test function out
vector <- c(0,3,4,5,0,3,4,8,0,9,8)
count_zeros(x=vector)
```

### Question 2:
```{r, message=FALSE, warning=FALSE}
# same thing but with subsetting
count_zeros2 <- function(x) {
  return(length(x[x==0]))
}

# test function out
vector <- c(0,3,4,5,0,3,4,8,0,9,8)
count_zeros2(x=vector)
```

### Question 3:
```{r, message=FALSE, warning=FALSE}
# create function
make_matrix <- function(rows,columns) {
  result <- matrix(NA, nrow = rows, ncol = columns)
  for (i in 1:nrow(result)) {
    for (j in 1:ncol(result)) {
      result[i,j] <- i*j
    }
  }
  return(result)
}

# test function out
make_matrix(6,7)
```

### Question 4a:
```{r, message=FALSE, warning=FALSE}
nGroup <- 3 # number of groups
nName <- c("Group1","Group2", "Group3") # names of groups
nSize <- c(30,30,30) # number of observations
nMean <- c(20,40,60) # mean of each group

response <- c(rnorm(n=nSize[1],mean=nMean[1]),
            rnorm(n=nSize[2],mean=nMean[2]),
            rnorm(n=nSize[3],mean=nMean[3]))
groups <- rep(nName,nSize)
my_data <- data.frame(groups,response)
my_data
```

### Question 4b:
```{r, message=FALSE, warning=FALSE}
shuffleMeans <- function(data=my_data){
  shuffled <- sample(data$response)
  data$shuffled <- shuffled
  result <- c(mean(shuffled[data$groups=="Group1"]),mean(shuffled[data$groups=="Group2"]),mean(shuffled[data$groups=="Group3"]))
  return(result)
}

shuffleMeans()
```

### Question 4c:
```{r, message=FALSE, warning=FALSE}
data100replicates <- data.frame(Replicate = 1:100, Group1 = 100, Group2 = 100, Group3 = 100)

for (i in 1:100){
  new_means <- shuffleMeans(my_data)
  data100replicates[i, 2:4] <- new_means
}
print(data100replicates)
```

### Question 4d
```{r, message=FALSE, warning=FALSE}
# plot for group 1
qplot(data=data100replicates,x=Group1,geom="histogram") +
  labs(x = "Mean", y = "Frequency", title = "Group 1")

# plot for group 2
qplot(data=data100replicates,x=Group2,geom="histogram") +
  labs(x = "Mean", y = "Frequency", title = "Group 2")

# plot for group 3
qplot(data=data100replicates,x=Group3,geom="histogram") +
  labs(x = "Mean", y = "Frequency", title = "Group 3")
```
The distributions of the reshuffled means look quite similar across groups. The distributions of the original means looked quite different across groups. This makes sense because we reshuffled the data across groups, which reduced the differences in our response that we originally saw between the groups.