---
title: "Homework_8"
author: "Kyle Grasso"
date: "2024-04-10"
output: html_document
---
### Question 1:
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(datasets)
str(iris)
```
##### There are 150 observations and 5 variables in the iris dataset.

### Question 2:
```{r, message=FALSE, warning=FALSE}
iris1 <- filter(iris, Species != "setosa" & Sepal.Length > 6 & Sepal.Width > 2.5)
str(iris1)
```
##### There are now only 56 observations and 5 variables in this updated dataset

### Question 3:
```{r, message=FALSE, warning=FALSE}
iris2 <- select (iris1, Species, Sepal.Length, Sepal.Width)
str(iris2)
```
##### There are now 56 observations, but only the 3 variables that we selected.

### Question 4:
```{r, message=FALSE, warning=FALSE}
iris3 <- arrange(iris2, by = desc(Sepal.Length))
head(iris3)
```

### Question 5:
```{r, message=FALSE, warning=FALSE}
iris4 <- mutate(iris3, Sepal.Area = Sepal.Length*Sepal.Width)
str(iris4)
```
##### There are now 56 observations and 4 variables because we just added Sepal.Area as a variable.

### Question 6:
```{r, message=FALSE, warning=FALSE}
iris5 <- summarize(iris4, MeanSepalLength = mean(Sepal.Length), MeanSepalWidth = mean(Sepal.Width), SampleSize = n())
print(iris5)
```

### Question 7:
```{r, message=FALSE, warning=FALSE}
iris6 <- iris4 %>% group_by(Species) %>% summarize(MeanSepalLength = mean(Sepal.Length), MeanSepalWidth = mean(Sepal.Width), SampleSize = n())
print(iris6)
```

### Question 8:
```{r, message=FALSE, warning=FALSE}
irisFinal <- iris %>% 
  filter(Species != "setosa" & Sepal.Length > 6 & Sepal.Width > 2.5) %>% 
  select (Species, Sepal.Length, Sepal.Width) %>%
  arrange(by = desc(Sepal.Length)) %>%
  mutate(Sepal.Area = Sepal.Length*Sepal.Width) %>%
  group_by(Species) %>% 
  summarize(MeanSepalLength = mean(Sepal.Length), MeanSepalWidth = mean(Sepal.Width), SampleSize = n())

print(irisFinal)
```

### Question 9:
```{r, message=FALSE, warning=FALSE}
irisLonger <- iris %>% pivot_longer(cols=c(Sepal.Length,Sepal.Width,Petal.Length,Petal.Width), names_to = "Measure", values_to = "Value")
head(irisLonger)
```