---
title: "Homework_12"
author: "Kyle Grasso"
date: "2024-04-17"
output: html_document
---
### Loading libraries
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(wesanderson)
library(scales)
library(patchwork)
```

### Assigning and looking at structure of my data
I will focus on the Sepal.Length variable.
```{r, message=FALSE, warning=FALSE}
data <- iris
str(data)
```

### Basic Box Plot of Sepal Length
```{r, message=FALSE, warning=FALSE}
ggplot(data) + aes(x=Sepal.Length,y=Species,fill=Species) +
  geom_boxplot()
```

### Add formal title and axis labels to box plot
```{r, message=FALSE, warning=FALSE}
ggplot(data) + aes(x=Sepal.Length,y=Species,fill=Species) + geom_boxplot() + labs(title = "Boxplot", x = "Sepal Length (cm)", y="Plant Species") + theme(plot.title=element_text(hjust=0.5)) + theme(legend.position = "none")
```

### Add points, change theme, and create final box plot
```{r, message=FALSE, warning=FALSE}
my_colors <- c(wes_palettes$GrandBudapest2)

p1 <- ggplot(data) + aes(x=Sepal.Length,y=Species,fill=Species) + geom_boxplot() + labs(title = "Boxplot", x = "Sepal Length (cm)", y="Plant Species") + theme(plot.title=element_text(hjust=0.5)) + geom_point(position=position_jitter(width=0.7,height=0.2),color="grey60",size=3) + theme(legend.position = "none") + scale_fill_manual(values = my_colors)

p1
```

### Basic Density Plot of Sepal Length
```{r, message=FALSE, warning=FALSE}
ggplot(data) + aes(x=Sepal.Length,fill=Species) +
  geom_density()
```

### Create final density plot
```{r, message=FALSE, warning=FALSE}
p2 <- ggplot(data) + aes(x=Sepal.Length,fill=Species) + geom_density() + labs(title = "Density Plot", x = "Sepal Length (cm)", y="Frequency") + theme(plot.title=element_text(hjust=0.5)) + scale_fill_manual(values = my_colors)

p2
```

### Create final jitter plot
```{r, message=FALSE, warning=FALSE}
p3 <- ggplot(data) + aes(x=Species,y=Sepal.Length,color=Species) + geom_jitter() + labs(title = "Jitter Plot", x = "Plant Species", y="Sepal Length (cm)") + theme(plot.title=element_text(hjust=0.5))

p3
```

### Combine all the plots into one final presentable plot
```{r, message=FALSE, warning=FALSE}
final_plot <- (p1|p2)/(p3)+plot_annotation(caption = "a. Boxplot of Sepal Length by Species
b. Density Plot of Sepal Length by Species
c. Jitter Plot of Sepal Length by Species", tag_levels = "a")

final_plot
```